---
title: "Penguins EDA"
format:
  html:
    code-fold: true
---

Embark on a journey to the windswept shores and icy waters of the Palmer Archipelago, where the resilient penguins waddle the line between adorability and awe-inspiring adaptation. As ambassadors of the Antarctic, these penguins offer a unique glimpse into the power of evolution and the rich biodiversity of our planet's southernmost frontiers. Here, we dive into the data beneath the feathers, revealing a narrative written over millennia by ice, ocean, and the indomitable will to survive. Join us as we decode the secrets of their sizes, the nuances of their masses, and the very flippers that ferry them through frigid seas. This is not just a tale of birds in tuxedos but a saga of life thriving against the stark backdrop of the Earth's harshest habitat.

## Penguin Size and Mass by Sex and Species

``` {.r .cell-code}
library(palmerpenguins)
library(dplyr)
library(ggplot2)

df <- palmerpenguins::penguins
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(palmerpenguins)
library(dplyr)
library(ggplot2)

df <- palmerpenguins::penguins

```

``` {.r .cell-code}
df %>%
  group_by(species, sex) %>%
  summarise(
    across(
      where(is.numeric), 
      \(x) mean(x, na.rm = TRUE)
      )
    ) %>%
  knitr::kable()
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df %>%
  group_by(species, sex) %>%
  summarise(
    across(
      where(is.numeric), 
      \(x) mean(x, na.rm = TRUE)
      )
    ) %>%
  knitr::kable()
```

```{}
```

## Penguin Size vs Mass by Species

The scatter plot above depicts a clear positive correlation between the bill length and body mass of penguins, segmented by species. Each species exhibits a distinct regression line, indicating unique growth patterns and physical development trends. The Gentoo penguins, represented in blue, display a steeper slope, reflecting a more pronounced increase in body mass relative to bill length compared to the Adelie and Chinstrap species. These differences highlight the varied ecological roles and feeding behaviors among the species, possibly influenced by their respective habitats.

``` {.r .cell-code}
df %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

## **Flipper Length by Different Species**

A histogram unfurls the story of penguin flippers - the aquatic wings that propel these birds through their briny realm. Here, we see the spread of flipper lengths that underpins each species' swimming prowess. The Gentoos, cloaked in an azure reminiscent of deep waters, boast a range suggesting diverse growth patterns and, perhaps, a versatility in their underwater escapades. Contrastingly, the Adelies' flipper length variation is as singular as their black and white plumage, implying a uniformity that might be key to their social and foraging behaviors.

```{r}
library(ggplot2)

df <- palmerpenguins::penguins
df <- na.omit(df) 

flipper_plot <- ggplot(df, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram(bins = 30, alpha = 0.7) +
  facet_wrap(~species) +
  labs(title = "Flipper Lengths by Species", x = "Flipper Length (mm)", y = "Count")

flipper_plot
```

## Male vs Female Penguins by Species

This section of the analysis brings forward a comparative overview of the mean physical characteristics of three penguin species seperated by sex. By summarizing these key metrics, we gain insight into the sexual dimorphism and inter-species differences that may be crucial for their adaptation to specific ecological niches. Noticeably, the Gentoo penguins generally show larger dimensions, suggesting a possible evolutionary advantage for their particular foraging and survival strategies.

```{r}
gender_plot <- ggplot(df, aes(x = flipper_length_mm, y = body_mass_g, color = sex)) +
  geom_point() +
  facet_wrap(~species) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Male vs Female Penguins by Species", x = "Flipper Length (mm)", y = "Body Mass (g)")

gender_plot

```

```{r}
#install.packages("DBI")
```

```{r}
#con <- DBI::dbConnect(
#  duckdb::duckdb(), 
#  dbdir = "my-db.duckdb")
#df <- dplyr::tbl(con, "penguins")
```

```{r}
#DBI::dbDisconnect(con, shutdown = TRUE )

```

As our exploration comes to a close, we leave the world of the Adelie, Chinstrap, and Gentoo penguins with newfound respect and fascination. From the stately stride of a Gentoo to the intricate interplay of size and survival that characterizes these species, our analysis has brought to light the remarkable resilience and adaptability of these Antarctic inhabitants. The data speaks volumes, not only of ecological dynamics but also of a larger story that connects us all to the rhythms of the natural world. It's a narrative that continues beyond the borders of this website, in the vast, open library of life where every creature pens its own chapter. As stewards of these narratives, we're reminded of the delicate balance that life maintains and the imperative role we play in preserving it. Thank you for sharing in this journey of discovery, and may your curiosity about the natural world continue to lead you to new horizons.

```{r}
library(rmarkdown)
```
